Quantitative Auditing of Fairness Measures with Differentially Private Synthetic Data

### 2024-12-05 paper/paper.tex
related work

TODO: add more connections to our work

Fairness in machine learning has become a crucial area of research\cite{barocas2023fairness}. Various fairness measures have been proposed to quantify\cite{yeh2024analyzing} the fairness of AI models\cite{pessach2022review,corbett2017algorithmic,vzliobaite2017measuring,hardt2016equality,corbett2017algorithmic,berk2021fairness,chouldechova2017fair,kleinberg2016inherent}.

Auditing of AI models is an important step in ensuring that these models are fair\cite{ferrara2023fairness}. Different tools have been developed to check the fairness of AI models\cite{yuan2024ensuring,saleiro2018aequitas,bellamy2019ai,bird2020fairlearn}.

@@

Differential Privacy has risen to become the standard of privacy protection in many data analyses\cite{jiang2021applications}. It offers a formal framework for quantifying privacy guarantees when releasing information derived from sensitive data\cite{dwork2006calibrating,dwork2014algorithmic}.

Different flavors of Differential Privacy have been proposed over the years\cite{desfontaines2019sok}, such as Gaussian Differential Privacy\cite{dong2022gaussian}, Pufferfish Differential Privacy\cite{kifer2012rigorous}, Bayesian Differential Privacy\cite{triastcyn2020bayesian}, and R\'enyi Differential Privacy\cite{mironov2017renyi}.

@@

The generation of synthetic data\cite{raghunathan2021synthetic,lu2023machine} under Differential Privacy constraints allows for data sharing without compromising privacy\cite{tao2021benchmarking}. There have been several techniques developed for Differentially Private synthetic data generation\cite{rosenblatt2020differentially,fan2020relational,bowen2019comparative,bowen2020comparative,arnold2020really,xu2019modeling} in recent years.

TODO: some works aim to change properties of og data. we want to preserve. so we dont consider this case

There are also works that aim to introduce bias to synthetic data\cite{jiang2024synthetic,baumann2023bias}.

While helpful, there are some technical limitations\cite{stadler2022synthetic,cheng2021can,ganev2022robin,wyllie2024fairness} and ethical concerns\cite{whitney2024real} with the use of synthetic data.

@@

% \begin{table}[h]
% \caption{Example marginals.}
% \label{tab:marginal-example}
% \centering
% \begin{subtable}[t]{0.3\linewidth}
%   \caption{Fairness measures experiment results of the adult dataset. Average of differences is 0.0431.}
%   \label{tab:adult_score}
%   \begin{tabular}{ccccc}
%   \toprule
%   \textbf{Measure} & \textbf{Original} & \textbf{Synthetic} & \textbf{Difference} \\
%   \midrule
%   Demographic Parity  & 0.172 & 0.104 & 0.067 \\
%   Accuracy Eqaulity   & 0.047 & 0.117 & 0.069 \\
%   Equalized Odds 1    & 0.057 & 0.079 & 0.021 \\
%   Equalized Odds 2    & 0.166 & 0.122 & 0.044 \\
%   Accuracy Equality 1 & 0.100 & 0.132 & 0.031 \\
%   Accuracy Equality 2 & 0.119 & 0.146 & 0.027 \\
%   \bottomrule
%   \end{tabular}
% \end{subtable}
% \begin{subtable}[t]{0.3\linewidth}
%   \centering
%   \caption{Fairness measures experiment results of the COMPAS dataset. Average of differences is 0.031.}
%   \label{tab:diabetes_score}
%   \begin{tabular}{cccc}
%   \toprule
%   \textbf{Measure} & \textbf{Original} & \textbf{Synthetic} & \textbf{Difference} \\
%   \midrule
%   Demographic Parity  & 0.013 & 0.015 & 0.002 \\
%   Accuracy Eqaulity   & 0.007 & 0.009 & 0.002 \\
%   Equalized Odds 1    & 0.000 & 0.003 & 0.003 \\
%   Equalized Odds 2    & 0.008 & 0.106 & 0.097 \\
%   Accuracy Equality 1 & 0.013 & 0.097 & 0.084 \\
%   Accuracy Equality 2 & 0.021 & 0.020 & 0.001 \\
%   \bottomrule
%   \end{tabular}
% \end{subtable}
% \end{table}

@@

% TODO: mention model fit time? yes. do experiments!

@@

% %  TODO: rewrite this

@@

%  TODO: rewrite this

@@

% TODO: expand all defs and count the nums of avail vars

>>>>

% fairness

Fairness in machine learning has become a crucial area of research\cite{barocas2023fairness}. Various fairness measures have been proposed to quantify\cite{yeh2024analyzing} the fairness of AI models\cite{pessach2022review,corbett2017algorithmic,vzliobaite2017measuring,hardt2016equality,corbett2017algorithmic,berk2021fairness,chouldechova2017fair,kleinberg2016inherent}. We consider these fairness measures in this work.

Auditing of AI models is an important step in ensuring that these models are fair\cite{ferrara2023fairness}. Different tools have been developed to check the fairness of AI models\cite{saleiro2018aequitas,bellamy2019ai,bird2020fairlearn}. We use a tool produced from our previous work\cite{yuan2024ensuring}.

@@

Differential Privacy has risen to become the standard of privacy protection in many data analyses\cite{jiang2021applications}. It offers a formal framework for quantifying privacy guarantees when releasing information derived from sensitive data\cite{dwork2006calibrating,dwork2014algorithmic}. We use Differential Privacy in this work to protect the privacy of individuals in the data in our auditing framework.

Different flavors of Differential Privacy have been proposed over the years\cite{desfontaines2019sok}, such as Gaussian Differential Privacy\cite{dong2022gaussian}, Pufferfish Differential Privacy\cite{kifer2012rigorous}, Bayesian Differential Privacy\cite{triastcyn2020bayesian}, and R\'enyi Differential Privacy\cite{mironov2017renyi}. We consider R\'enyi Differential Privacy specifically in this work.

@@

The generation of synthetic data\cite{raghunathan2021synthetic,lu2023machine} under Differential Privacy constraints allows for data sharing without compromising privacy\cite{tao2021benchmarking}. There have been several techniques developed for Differentially Private synthetic data generation\cite{rosenblatt2020differentially,fan2020relational,bowen2019comparative,bowen2020comparative,arnold2020really,xu2019modeling} in recent years. We use the technique that won the 2018 NIST Differential Privacy Synthetic Data Challenge competition\cite{mckenna2021winning}.

There have also been some works that aim to alter the properties of synthetic data, such as introducing bias to them\cite{jiang2024synthetic,baumann2023bias}. In this work, we aim to preserve the properties of the original data. Thus, we do not consider these works.

While helpful, there are some technical limitations\cite{stadler2022synthetic,cheng2021can,ganev2022robin,wyllie2024fairness} and ethical concerns\cite{whitney2024real} with the use of synthetic data. Synthetic data may not always preserve properties of real data. This may cause inaccuracies in downstream tasks such as fairness evaluation. Our work examines these limitations.

@@

@@

@@

@@

@@

% expand all defs and count the nums of avail vars
